---
title: "Discards Proportion Data"
author: "Camila Vargas"
date: "3/27/2022"
output: html_document
---

Can we estimate how much Blue Shark gets disccarded at sea based on the Third Assesment of Discards data?

Things to consider:
- The MLE model used for the third assessment calculates discards per fishery and not per species. This means we can calculate discards per flag and gear but not at a species level.
- On the other hand, we know the percentage of each FAO species going into each fishery (this and the point above are based on average landing data from 2010 to 2014)
Assuming that the proportion of discard/landing for each fishery is the same for each species considered in a fishery, I think we could calculate the discard per species(by flag and gear type) using the fishery ration of discard/landing. I'm not sure if this is a valid assumption though. 
- With this method, we can only calculate discards for the species considered in the Third assessment. This probably includes most blue sharks; however, there might be discrepancies with recent data. 

## Set-up
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(readxl)
library(janitor)

```


## Read data
```{r}
discard_data_url <- "https://raw.githubusercontent.com/bkuczenski/scoping-gear-losses/main/data/perez-roda_discards/fao_elg_landdisc.csv"

disc_data <- read_csv(discard_data_url) %>% 
  clean_names()

## How much of each species goes into each fishery
fishery_table <- read_excel(here::here("data/datarepository.xlsm"), sheet = "FisheryTable") %>% 
  clean_names() %>% 
  select(id, country, fishery_name = name, fao_area, ocean, depth, location, target, target_species, vessel_length, gear, fish_stat)

## Average landings (2010-2014) of each species by country
fish_stat <- read_excel(here::here("data/datarepository.xlsm"), sheet = "FishStat") %>% 
  clean_names() %>% 
  select(fish_stat_id, country,fishing_area_name,fishing_area_code, spp_common, spp_scientific, spp_alpha, land_2014, land_avg_10_14)

```


## Idenfifying Fish Stat IDs

Looking into Fishery Table to identify all the FishStat IDs and it's percentage that go into one fishery.

```{r}
fish_stat_codes <- fishery_table %>% 
  mutate(fish_stat_vec = str_split(fish_stat, ";")) %>% 
  unnest(fish_stat_vec) %>% 
  filter(fish_stat_vec != "") %>% 
  mutate(position = rep(c("fish_stat_id","pct"), times = n()/2),
         category = rep(1:(n()/2), each = 2)) %>% 
  pivot_wider(names_from = position,
              values_from = fish_stat_vec) %>% 
  select(-fish_stat, - category) %>% 
  mutate(fish_stat_id = as.numeric(fish_stat_id),
         country = case_when(country == "cabo verde" ~ "Cabo Verde",
                             T ~ country)) %>% 
  arrange(fish_stat_id)
##Check all species sum to 100%
check_pct <- fish_stat_codes %>% 
  group_by(fish_stat_id) %>% 
  mutate(pct = as.numeric(pct)) %>% 
  summarise(pct_total = sum(pct)) %>% 
  filter(pct_total < 100)
##stat id 2786 (Artisanal shrimp from El Salvador) sums 10 probably its a typo mistake
```

## Matching 

Note: not all species in fish stats are in the Fishery Table.

```{r}
fish_stat_allocated <- fish_stat_codes %>%
  left_join(fish_stat, by = c("fish_stat_id", "country")) %>%
  mutate_at(c("spp_common", "spp_alpha"), str_to_lower) %>%
  filter(!fishing_area_name %in% c("Africa - Inland waters", "America, South - Inland waters", "Former USSR area - Inland waters", "Oceania - Inland waters", "America, North - Inland waters", "Asia - Inland waters", "Europe - Inland waters")) %>%
  mutate(pct = as.numeric(pct),
         catch_amount = land_avg_10_14 * pct / 100) %>%
  select( -ocean, -target_species, -spp_scientific)
```


